---
output: github_document
---


# Analyzing the flights data

> My cousin is getting married in Kansas City in May and I need to book flights from Baltimore to Kansas City. I want to decide which airline to pick, which weekday to depart, and what time of day to leave so that so that the I have the least chance of my plane being delayed. Which airline should I pick, which day of the week should I depart on, and around when during the day should I depart?

Now that we've [downloaded the data](https://github.com/simonpcouch/gwc-bmore-2022/blob/main/source/grad_data.md), we can analyze it to get at these questions.

First loading the packages I'll need,

```{r, message = FALSE, warning = FALSE}
# general tools to work with data
library(tidyverse)
library(lubridate)
```

Now, loading the data,

```{r}
load("source/flights.Rda")
```

Taking a glimpse at the data,

```{r}
# take a look by column
glimpse(flights)

# take a look by rows
head(flights)
```

Each row is a flight that left BWI (note: we don't have cancelled ones), and each column is some piece of information about that flights departure: time, date, departure delay, destination, etc.

### Which airline should I pick?

Note that there's a column in our data called `carrier`--that's our airline.

```{r}
head(flights$carrier)
```

There's also a column called `dep_delay` that gives us our departure delay for a given flight.

```{r}
head(flights$dep_delay)
```

We'll use some tools from the tidyverse to group by the airline and then take the average flight delay `dep_delay` for that airline. Then, we can visualize our results.

```{r}
mean_delay_by_carrier <-
  flights %>%
    group_by(carrier) %>%
    summarize(mean_delay = mean(dep_delay, na.rm = TRUE),
              n = n()) %>%
  arrange(mean_delay)

mean_delay_by_carrier
```

Okay—so, maybe I should pick United Air Lines! This is just an average, though—how much variation is there in departure delay?

```{r}
flights %>%
  filter(carrier == "UA") %>%
  ggplot() +
    geom_histogram(bins = 20) +
    aes(x = dep_delay)
```

Looks like, most of the time, United Air Lines flights left just before the scheduled time or right on time. When they left late, though, they left _late_.

### Which day of the week should I depart on?

This one's a bit tricky. In this data, we have the date that each flight leaves one, but we'll have to match that up with a calendar to get the weekday.

```{r}
flights <- flights %>%
  # make the `weekday` column
  mutate(weekday = wday(time_hour, label = TRUE))

head(flights$weekday)
```

Now that we have the weekday, this problem looks a lot like the one above,

```{r}
mean_delay_by_weekday <-
  flights %>%
    group_by(weekday) %>%
    summarize(mean_delay = mean(dep_delay, na.rm = TRUE),
              n = n()) %>%
  arrange(mean_delay)

mean_delay_by_weekday
```

Visualizing these results,

```{r}
mean_delay_by_weekday %>%
  ggplot() +
  aes(x = weekday, y = mean_delay) +
  geom_col()
```



Maybe Tuesday is the day to head out!

### Around when during the day should I depart?

```{r}
mean_delay_by_tod <-
  flights %>%
    mutate(tod = case_when(
      hour >= 1 & hour <= 6 ~ "early morning",
      hour >= 7 & hour <= 12 ~ "morning",
      hour >= 13 & hour <= 18 ~ "afternoon",
      hour >= 19 & hour <= 24 ~ "evening",
    )) %>%
    group_by(tod) %>%
    summarize(mean_delay = mean(dep_delay, na.rm = TRUE),
              n = n()) %>%
  arrange(mean_delay)

mean_delay_by_tod
```

Oh, interesting! The later in the day a flight leaves, the more likely it is to be delayed. Makes sense, I guess? Airlines can troubleshoot issues overnight and then have a "blank slate" each morning.

### Concluding thoughts

We've used data to help us inform our decision on which flights to take out of BWI. What things might influence how correct our analysis is?

Some possibilities:

* I looked at airline, weekday, and time of day _separately_. Do these things relate to each other in their influence on departure delays? That is, does the time of day matter more for delays on a weekday than a weekend? Are some airlines better at overnight flights than others?
* Delays might be caused by different factors in 2021 than in 2022. Supply chain issues, the financial wellness of different airlines, etc.
* I assumed that a flight occurring in May was a more important similarity than a flight being recent. Should I have used the most recent data possible?
* I only looked at flights that _left BWI_, and did not also require that the flight _arrived in Kansas City_. Maybe certain airlines are better at flying to certain places!
* We didn't consider flights that were cancelled, as they weren't included in the day. Maybe it's better for a flight to be delayed than cancelled, and delays mean reselience to cancellation!
